---
title: "Portfolio5"
author: "Conor Lacey"
date: "2023-03-31"
output: github_document
---
## Libraries
```{r libraries}
suppressWarnings(library(tidyverse))
suppressWarnings(library(dmacs))
source("functions2.R")
```


## Introduction

In this portfolio I will be running simulations to explore the nature of dMACS_Shrunk in comparison to the original dMACS measure. 

To do this we will be running conditions on sample size, prior probabilities, effect sizes, and whether or not the mean eta value (i.e., the latent variable) is equal across groups. To be more specific our sample size conditions will include sizes of 250, 500, and 1000. Our prior probabilities will include prior probabilities of the null being true of .25, .5, and .75. The effect size conditions will include effect sizes of 0.1, 0.5, and 0.9. Lastly, the mean eta value conditions will include one condition where they are exactly equal and another where there is a mean difference of 1. 

In total, we will have 54 conditions. This is quite a lot and I want to calculate these conditions 500 times! Therefore, the most intuitive way I believe I can organize this is by making a 500 X 54 data frame. Make sense? 

## Data Frames

Alright let's do this one step at a time. First, let's create the data frame. 

```{r data frames}
set.seed(9420)

names <- c(1:54) %>% as.character()

dat_dMACS <- matrix(nrow = 500, ncol = 54) %>% data.frame() %>% 
  setNames(c(names))
dat_dMACS_S <- matrix(nrow = 500, ncol = 54) %>% data.frame() %>% 
  setNames(c(names))
```

## For Loop Simulations

To conduct these simulations I will do a series of if statement wrapped in a for loop. This loop should calculate a dMACS and dMACS_Shrunk estimate under each of the 54 conditions and loop back to do this again 500 times.

```{for loops sims}
for (i in 1:500){
  for (cond in 1:54){
    if (cond %in% c(1,4,7,10,13,16,19,22,25,28,31,34,37,40,43,46,49,52)){
      N.R <- 250
      N.F <- 250
    }
    if (cond %in% c(2,5,8,11,14,17,20,23,26,29,32,35,38,41,44,47,50,53)){
      N.R <- 500
      N.F <- 500
    }
    if (cond %in% c(3,6,9,12,15,18,21,24,27,30,33,36,39,42,45,48,51,54)){
      N.R <- 1000
      N.F <- 1000
    }
    if (cond %in% c(1,2,3,10,11,12,19,20,21)){ # dMACS 0.1 mean equal
      load.R <- .6 #loading for the reference group
      load.F <- .634 #loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.05 #intercept for focal group
    }
    if (cond %in% c(28,29,30,37,38,39,46,47,48)){ #dMACS 0.1 mean unequal
      load.R <- .6 #loading for the reference group
      load.F <- .61 #loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.05 #intercept for focal group
    }
    if (cond %in% c(4,5,6,13,14,15,22,23,24)){ #dMACS 0.2 mean equal
      load.R <- .6 #loading for the reference group
      load.F <- .61 #loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.3 #intercept for focal group
    }
    if (cond %in% c(31,32,33,40,41,42,49,50,51)){ #dMACS 0.2 mean unequal
      load.R <- .6 #loading for the reference group
      load.F <- .61 #loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.29 #intercept for focal group
    }
    if (cond %in% c(7,8,9,16,17,18,25,26,27)){ #dMACS 0.9 mean equal
      load.R <- .6 #loading for the reference group
      load.F <- .606#loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.55 #intercept for focal group
    }
    if (cond %in% c(34,35,36,43,44,45,52,53,54)){ #dMACS 0.9 mean unequal
      load.R <- .6 #loading for the reference group
      load.F <- .595 #loading for focal group
      int.R <- 0 #intercept for reference group
      int.F <- 0.55 #intercept for focal group
    }
    if (cond %in% c(1:9,28:36)){
      priorPH0  <- 0.25
    }
    if (cond %in% c(10:18,37:45)){
      priorPH0 <- 0.5
    }
    if (cond %in% c(19:27,46:54)){
      priorPH0 <- 0.75
    }
    if (cond %in% c(1:27)){ #mean equal
      mean.F <- 0
    }
    if (cond %in% c(28:54)){ #mean unequal
      mean.F <- 1
    }
    
    mean.R <- 0
    
    
    sd.R <- 1
    sd.F <- 1
    
    #Set up latent variable
    eta.R <- rnorm(N.R, mean.R, sd.R)
    eta.F <- rnorm(N.F, mean.F, sd.F)
    
    y.R <- load.R*eta.R + int.R #model-implied scores for reference group
    y.F <- load.F*eta.F+ int.F #model-implied scores for focal group
    
    datR <- data.frame(eta = eta.R, y = y.R, group = "R")
    datF <- data.frame(eta = eta.F, y = y.F, group = "F")
    dat  <- bind_rows(datR,datF)
    
    dat %>% ggplot(aes(x = eta, y = y, color = group)) +
      geom_line(linewidth = 1) +
      labs(x = "Eta",
           y = "Estimated Response") +
      geom_line(data = datF,
                aes(x = eta, y = dnorm(eta, mean(eta), sd(eta)), color = NULL),
                linetype = "dashed") +
      xlim(-3,3) +
      ylim(0,1.2)
    
    sd.pooled.test.num <- ((length(y.R)-1)*sd(y.R)) + ((length(y.F)-1)*sd(y.F))
    sd.pooled.test.denom <- (length(y.R) - 1) + (length(y.F) - 1)
    sd.pooled.test <- sd.pooled.test.num/sd.pooled.test.denom
    
    #dMACS
    dMACS <- item_dmacs(load.R, #loading, R
                        load.F, #loading, F
                        int.R, #intercept, R
                        int.F, #intercept, F
                        mean(eta.F), #mean (both groups)
                        var(eta.F), #variance (both groups, but it would be the focal group if they didn't have the same variance)
                        sd.pooled.test #pooled standard deviation (again, this is basically arbitrary)
    )
    
    dat_dMACS[i,cond] <- dMACS
    
    #dMACS_Shrunk 
    sigmaSlab <- 1
    
    ybarExpl <- get.dMACS(y.R, y.F, eta.R, eta.F)
    ybarExpl
    
    nExpl <- length(y.R)
    
    upMAExpl <- updatePar(priorPH0, sigmaSlab, nExpl, ybarExpl[1L])
    ciMAExpl <- postStat(upMAExpl)
    
    tbExplicit <- data.frame(t(c(upMAExpl, ciMAExpl)))
    names(tbExplicit) <- c("ph0", "mu1", "sd1", "Lower", "Upper", "modelAveraged")
    
    dMACS_Shrunk <- (1-tbExplicit[1])*tbExplicit[2]
    
    dat_dMACS_S[i,cond]<- dMACS_Shrunk
  }
}
```

## Output
Yay! Simulations done. Now I will save the outputted data files.

```{Output}
write.csv(dat_dMACS, file = "dat_dMACS.csv")
write.csv(dat_dMACS_S, file = "dat_dMACS_S.csv")
```



